Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods

### 2022-09-25 inst/surfing_jsa_rebuttal_minorcomments.tex

Page 7/Figure 2: The authors suggest there is a loose positive
correlation between maximum likelihood strength and total points,
which is undoubtedly true. However, the x-axis on Figure 2 itself has
been distorted, which makes the trend appear more linear than it
really is. I'd prefer to see the x-axis undistorted, unless the
authors are explicitly trying to make a point about the trend being
log-linear, in which case I'd like to see this briefly discussed.

\textcolor{blue}{Horizontal axis now linear; no other change except
  for tiny movement of labels so they do not obscure one another.  We
  are mortified to see that the linearity is more apparent without the
  log-transform.}

Page 7: The authors suggest we may reject the hypothesis that the
strength of the entity exceeds about 0.025. Does this assertation come
with a supporting p-value? It is unclear from Figure 4 how this
conclusion is reached.

\textcolor{blue}{This was a (devil-inspired) typo.  The value should
  read 0.015.  The reasoning was that at 0.015 the support was -2.  We
  have ddded a sentence of explanation on p7 and also the figure
  caption.}

Grammatical changes

Page 3: Change ``Further, we use a generalization due to (Luce,
1959)..." to ``Further, we use a generalization due to Luce
(1959)...".

Page 7: Change ``..., but in addition we may reject the
the proposal that..." to ``..., but in addition we may reject the
proposal that...".

### 2022-09-25 inst/surfing_jsa.tex
fixes #173

The Bradley-Terry model \citep{bradley1954} assigns non-negative
strengths $p_1,\ldots, p_n$ to each of $n$ competitors in such a way
that the probability of $i$ beating $j\neq i$ in pairwise competition
is $\frac{p_i}{p_i+p_j}$; it is conventional to normalize so that
$\sum p_i=1$.  Further, we use a generalization due to
\citep{luce1959}, in which the probability of competitor~$i$ winning
in a field of $\left\lbrace 1,\ldots, n\right\rbrace$ is
$\frac{p_i}{p_1+\cdots +p_n}$.  Noting that there is information in
the whole of the finishing order, and not just the first across the
line, we can follow \citet{plackett1975} and consider the runner-up to
be the winner among the remaining competitors, and so on down the
finishing order. Without loss of generality, if the order of finishing
were $1,2,3,4,5$ [we write $1\succ 2\succ 3\succ 4\succ5$], then a
suitable Plackett-Luce likelihood function would be

@@

To illustrate the concept, we show a simple likelihood function in
Table~\ref{probabilitybrazilian}.  The null would be that this reified
entity has zero strength.  We fail to reject this null (differential
support = 0.035; $p=0.79$) and conclude that there is no evidence to
suggest that Brazilians' performance is better on their preferred wave
types than one would expect by chance.  We may go further using the
method of support: not only is the maximum likelihood estimate for
Brazilian wavetype very small (about 0.002), but in addition we may
reject the the proposal that the strength of this entity exceeds about
0.015 (Figure~\ref{brazilian_wavetype_support}): at this value of the
strength or greater, we may gain two or more units of support by
moving to the evaluate~\citep{edwards1992}.

>>>>

The Bradley-Terry model \citep{bradley1954} assigns non-negative
strengths $p_1,\ldots, p_n$ to each of $n$ competitors in such a way
that the probability of $i$ beating $j\neq i$ in pairwise competition
is $\frac{p_i}{p_i+p_j}$; it is conventional to normalize so that
$\sum p_i=1$.  Further, we use a generalization due to
\citet{luce1959}, in which the probability of competitor~$i$ winning
in a field of $\left\lbrace 1,\ldots, n\right\rbrace$ is
$\frac{p_i}{p_1+\cdots +p_n}$.  Noting that there is information in
the whole of the finishing order, and not just the first across the
line, we can follow \citet{plackett1975} and consider the runner-up to
be the winner among the remaining competitors, and so on down the
finishing order. Without loss of generality, if the order of finishing
were $1,2,3,4,5$ [we write $1\succ 2\succ 3\succ 4\succ5$], then a
suitable Plackett-Luce likelihood function would be

@@

To illustrate the concept, we show a simple likelihood function in
Table~\ref{probabilitybrazilian}.  The null would be that this reified
entity has zero strength.  We fail to reject this null (differential
support = 0.035; $p=0.79$) and conclude that there is no evidence to
suggest that Brazilians' performance is better on their preferred wave
types than one would expect by chance.  We may go further using the
method of support: not only is the maximum likelihood estimate for
Brazilian wavetype very small (about 0.002), but in addition we may
reject the proposal that the strength of this entity exceeds about
0.015 (Figure~\ref{brazilian_wavetype_support}): at this value of the
strength or greater, we may gain two or more units of support by
moving to the evaluate~\citep{edwards1992}.
